Random Forests", OAs

### 2024-01-27 src/la_RF.Rmd
local minor changes
refined model OA
for bham presentation

```{r, eval=T}
firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>% 
  dplyr::select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
  rename(ladcd = GEOGRAPHY_CODE,
         firms = OBS_VALUE)

@@

@@

# cities
city.names <- c("London, UK", 
"Birmingham, UK",
"Glasgow, UK",
"Liverpool, UK",
"Bristol, UK",
"Manchester, UK",
"Sheffield, UK",
"Leeds, UK",
"Edinburgh, UK",
"Leicester, UK")

@@

# Join with distance and area
df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
  left_join(la, by = c("ladcd" = "LAD21CD")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,     # UNITS APPEAR WRONGLY IN M AND NOT IN KM
         density = n / area) %>% 
  relocate(density, .after = area)
sapply(df, function(x) sum(is.na(x)))
```

@@

# nearest city
help1 <- st_filter( la, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Edinburgh",
                    "Glasgow"))

@@

# This efficiently returns the name of the column with the shortest distance
dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:11]*-1)+1L])
help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
  rename(nearest.city.LAD21CD = LAD21CD) %>% 
  dplyr::select(ladcd, nearest.city.LAD21CD)

@@

df <- df %>% drop_na() %>% # THIS DROPS L999999
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(ladcd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("ladcd" = "ladcd"))

@@

df <- df %>% filter(year > 1996)
sapply(df, function(x) sum(is.na(x)))
```

@@

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df %>% filter(year > 1999),                     # FILTER BY YEAR FOR GROWTH
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

@@

start.time <- Sys.time()
model.all <- train(growth ~                                                     # REPLACE n WITH growth
                     #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
                     I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                   data = df %>% filter(year > 1999),                           # FILTER BY YEAR FOR GROWTH
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

@@

print(model.all)
# Random Forest 
# 
# 5984 samples
#   10 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 4718, 4690, 5040, 4718, 5055, 4704, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE        Rsquared   MAE       
#    2    variance    0.02194750  0.9015001  0.01162254
#    2    extratrees  0.02245492  0.8984940  0.01229071
#    6    variance    0.02208770  0.9025087  0.01175267
#    6    extratrees  0.02165214  0.9081808  0.01161281
#   10    variance    0.02447416  0.8859265  0.01326348
#   10    extratrees  0.02204464  0.9064730  0.01184945
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 6, splitrule = extratrees and min.node.size = 5.

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.slag", "spatial lag of web density",
                                               ifelse(rowname == "n.lag", "temporal lag of web density",
                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag  of web density",
                                                             ifelse(rowname == "n.nearest.city", "web density in nearest city", rowname)))))))) %>% 
  column_to_rownames()

path.out <- paste0(path, "/outputs/rf/varimp_LA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)

# ffsmodel_LLO <- ffs(predictors = df[, c("area", "dist", "dist.retail", "London", "south",  
#                                       "n.slag", "n.lag", "n.l.slag", "n.nearest.city", 
#                                                         "year")], 
#                     response = df$n, #%>% dplyr::select(n) %>% as.data.frame %>% mutate(n = as.numeric(n)),
#                     trControl = tc,
#                     method = "ranger",
#                     #na.action = na.omit,
#                     #importance = "impurity",
#                     metric="Rsquared",
#                     verbose=FALSE)
# ffsmodel_LLO
# ffsmodel_LLO$selectedvars
# plot_ffs(ffsmodel_LLO)
```

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
set.seed(831)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 6) # 4 feature + 2

@@

  model.all <- train(growth ~                                                   # REPLACE n WITH growth
                       #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                       I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                   data = df %>% filter(RGN21CD!=i), #& year > 1999),           # FILTER BY YEAR FOR GROWTH
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
}

stopCluster(cl)
end.time <- Sys.time() # ca. 2h
time.taken <- round(end.time - start.time,2)
time.taken
```

@@

    d <- paste0("region.predict.model.", ".on.", i)
    d <- cbind(get(d), df %>% filter(RGN21CD==i)) #& year > 1999))                      # FILTER BY YEAR
    d <- d %>% dplyr::select(RGN21CD, ladcd, year, growth, 1)                           # REPLACE n WITH growth
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(tested.on = i)
    df.predictions <- rbind(d, df.predictions)
}

@@

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$growth))    # REPLACE n WITH growth

@@

path.out <- paste0(path, "/outputs/rf/heatmap_LA.png")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "test.on") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
  arrange(Rsquared)

ggplot(aes(train.region, test.region, fill= Rsquared)) + 
  geom_tile() +
  coord_fixed() +
  labs(y = "", x = "") + #y: test.region, x: train.region
  scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
  theme(#legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust=1),
        panel.border = element_blank(), 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_rect(fill = "white"))
ggsave(path.out)

>>>>

```{r, eval=T}

# also saved locally in /data/raw/firm_data_backup.csv

firms <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_142_1.data.csv?geography=1811939329...1811939332,1811939334...1811939336,1811939338...1811939428,1811939436...1811939442,1811939768,1811939769,1811939443...1811939497,1811939499...1811939501,1811939503,1811939505...1811939507,1811939509...1811939517,1811939519,1811939520,1811939524...1811939570,1811939575...1811939599,1811939601...1811939628,1811939630...1811939634,1811939636...1811939647,1811939649,1811939655...1811939664,1811939667...1811939680,1811939682,1811939683,1811939685,1811939687...1811939704,1811939707,1811939708,1811939710,1811939712...1811939717,1811939719,1811939720,1811939722...1811939730,1811939757...1811939767&date=latestMINUS10&industry=37748736&employment_sizeband=0&legal_status=0&measures=20100") %>% 
  dplyr::select(GEOGRAPHY_CODE, OBS_VALUE) %>% #GEOGRAPHY_NAME, 
  rename(ladcd = GEOGRAPHY_CODE,
         firms = OBS_VALUE)

@@

# cities
cities.1 <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  slice_max(pop, n = 10)
cities.2 <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  filter(name == "Belfast")

cities.sf <- bind_rows(cities.1, cities.2) %>% 
  st_as_sf(coords = c("long", "lat"), crs = 4326)
```

@@

# cities
city.names <- c("London, UK", 
"Birmingham, UK",
"Glasgow, UK",
"Liverpool, UK",
"Bristol, UK",
"Manchester, UK",
"Sheffield, UK",
"Leeds, UK",
"Edinburgh, UK",
"Leicester, UK",
"Belfast, UK")

@@

# Join with distance and area
df <- df %>% left_join(dist, by = c("ladcd" = "ladcd")) %>% 
  left_join(la, by = c("ladcd" = "LAD21CD")) %>% 
  relocate(area, .after = n) %>% 
  mutate(area = area / 1000000,     # UNITS APPEAR WRONGLY IN M AND NOT IN KM
         density = n / area) %>% 
  relocate(density, .after = area)

df <- df %>% filter(!is.na(London))  # to drop 34 for L99999999 and M99999999

@@

#************** DOUBLE CHECK THE ORDER OF CITY NAMES BELOW **************

# nearest city
help1 <- st_filter( la, cities.sf) %>% as_tibble %>% dplyr::select(LAD21NM,LAD21CD) %>% 
  mutate(cities = c("Leicester",
                    "Bristol",
                    "Manchester",
                    "Liverpool",
                    "Sheffield",
                    "Birmingham",
                    "Leeds",
                    "London",
                    "Belfast",
                    "Edinburgh",
                    "Glasgow"))

@@

# This efficiently returns the name of the column with the shortest distance
dist <- dist %>% mutate(dist.city.name = names(.)[max.col(.[2:12]*-1)+1L])
help2 <- dist %>% left_join(help1, by = c("dist.city.name" = "cities")) %>% 
  rename(nearest.city.LAD21CD = LAD21CD) %>% 
  dplyr::select(ladcd, nearest.city.LAD21CD)

@@

df <- df %>% 
  group_by(year) %>% 
  mutate(n.slag = lag.listw(knn.l, n)) %>% 
  group_by(ladcd) %>% 
  mutate(n.l.slag = dplyr::lag(n.slag, n=1, order_by=year),
         n.lag = dplyr::lag(n, n=1, order_by = year)) %>% 
  left_join(help2, by = c("ladcd" = "ladcd"))

@@

# There are no lags for 1996, so it is removed
df <- df %>% filter(year > 1996)
sapply(df, function(x) sum(is.na(x)))
```

@@

```{r}
set.seed(123)
indices <- CreateSpacetimeFolds(df,# %>% filter(year > 1999),                   # FILTER BY YEAR FOR GROWTH
                                spacevar = "ladcd",
                                timevar = "year", 
                                k = 10)

@@

start.time <- Sys.time()
model.all <- train(n ~                                                          # REPLACE n WITH growth
                     #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + London.n.lag + n.nearest.city + year,
                     #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                     n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist, #n.nearest.retail.lag + dist.retail,
                   data = df,# %>% filter(year > 1999),                         # FILTER BY YEAR FOR GROWTH
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")

@@

print(model.all)

# [reproducible results]
# Random Forest 
# 
# 5984 samples
#    6 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 5055, 4690, 4732, 5070, 5040, 4718, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE        Rsquared   MAE       
#   2     variance    0.03227897  0.8023185  0.01991724
#   2     extratrees  0.03150004  0.8057960  0.01912348
#   4     variance    0.03271887  0.7985855  0.02025685
#   4     extratrees  0.03174791  0.8047090  0.01932473
#   6     variance    0.03358260  0.7879953  0.02080616
#   6     extratrees  0.03207532  0.8018495  0.01964516
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 2, splitrule = extratrees and min.node.size = 5.

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "n.London.lag", "London's web density, t-1", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.l.slag", "spatial and temporal lag of web density",
                                                             ifelse(rowname == "n.nearest.city.lag", "nearest city's web density, t-1", rowname)))))) %>% 
  column_to_rownames()

path.out <- paste0(path, "/outputs/rf/figures/varimp_LA.png")
ggplot(varimp_mars) + theme_minimal() + labs(x="") #+ theme(panel.background = element_rect(fill = "white"))
ggsave(path.out)
```

@@

#(8 is the number of tuning parameter, mtry for rf, here equal to ncol(iris)-2)
set.seed(831)
for(i in 1:10) seeds[[i]]<- sample.int(n=1000, 8) # 6 feature + 2

@@

  model.all <- train(n ~                                                        # REPLACE n WITH growth
                       #area + dist + dist.retail + London + south + n.slag + n.lag + n.l.slag + n.nearest.city + year,
                       #I(n.London.lag/dist) + I(n.nearest.city.lag/London) + n.l.slag + year,
                       n.London.lag + n.nearest.city.lag + n.l.slag + year + London + dist,
                   data = df %>% filter(RGN21CD!=i), #& year > 1999),           # FILTER BY YEAR FOR GROWTH
                   trControl = tc,
                   method = "ranger",
                   na.action = na.omit,
                   #preProc = c("center", "scale"),
                   importance = "impurity")
  assign(paste0("model.all",as.character(i)), model.all) # object name = region not in the training data
}

stopCluster(cl)
end.time <- Sys.time() 
time.taken <- round(end.time - start.time,2)
time.taken # ca. 4 minutes 
```

@@

    d <- paste0("region.predict.model.", ".on.", i)
    d <- cbind(get(d), df %>% filter(RGN21CD==i)) #& year > 1999))                      # FILTER BY YEAR
    d <- d %>% dplyr::select(RGN21CD, ladcd, year, n, 1)                                # REPLACE n WITH growth
    colnames(d)[5] <- "predictions"
    d <- d %>% mutate(tested.on = i)
    df.predictions <- rbind(d, df.predictions)
}

@@

# calculate metrics for every region pair 
rf.year.all.metrics <- lapply(pred.by.region.all.list, function(x) postResample(pred = x$predictions,
                                                                     obs = x$n))       # REPLACE n WITH growth

@@

path.out <- paste0(path, "/outputs/rf/figures/test_regions.csv")
rf.year.all.metrics %>% 
  as.data.frame() %>%
  rownames_to_column(var = "metrics") %>% 
  pivot_longer(!metrics, names_to = "test.on") %>% 
  pivot_wider(names_from = metrics, values_from = value) %>% 
  left_join(lookup.region, by = c("test.on" = "RGN21CD")) %>% 
  arrange(Rsquared) %>% 
  dplyr::select(RGN21NM, Rsquared) %>% 
  rename(Region = RGN21NM) %>% 
  write_csv(path.out)
  #ggplot(aes(x = 1, y = reorder(Region, Rsquared), fill = Rsquared)) +geom_tile()

# ggplot(aes(train.region, test.region, fill= Rsquared)) + 
#   geom_tile() +
#   coord_fixed() +
#   labs(y = "", x = "") + #y: test.region, x: train.region
#   scale_fill_gradientn(colours = terrain.colors(10), trans = 'reverse') +
#   theme(#legend.position = "none",
#         axis.text.x = element_text(angle = 45, hjust=1),
#         panel.border = element_blank(), 
#         panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(),
#         panel.background = element_rect(fill = "white"))
# ggsave(path.out)

  #summarise(r = range(Rsquared))
  #round(2) %>%
  #kable()
```

### 2024-01-27 src/oa_RF.Rmd
for bham presentation

# Random Forest 
# 
# 3716736 samples
#      10 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 2926910, 2926924, 3136005, 2926924, 3136005, 2926924, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE      Rsquared   MAE      
#    2    variance    3.865809  0.5809699  0.5313035
#    2    extratrees  3.730542  0.6168954  0.5884731
#    6    variance    3.761356  0.6063340  0.5366004
#    6    extratrees  3.612670  0.6273041  0.5331196
#   10    variance    3.929176  0.5775066  0.5449132
#   10    extratrees  3.617598  0.6278302  0.5403551
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 6, splitrule =
#  extratrees and min.node.size = 5.

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.slag", "spatial lag of web density",
                                               ifelse(rowname == "n.lag", "temporal lag of web density",
                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag of web density",
                                                             ifelse(rowname == "n.nearest.city", "web density in nearest city", rowname)))))))) %>% 
  column_to_rownames()

@@

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD==i), 
                                spacevar = "oa11cd",
                                timevar = "year", 
                                k = k)

@@

  file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
  save(model.all, file = file)  # file name = region not in the training data
  rm(model.all)

>>>>

# Random Forest 
# 
# 3716736 samples
#       8 predictor
# 
# No pre-processing
# Resampling: Cross-Validated (10 fold) 
# Summary of sample sizes: 2926910, 2926924, 3136005, 2926924, 3136005, 2926924, ... 
# Resampling results across tuning parameters:
# 
#   mtry  splitrule   RMSE      Rsquared   MAE     
#   2     variance    4.999920  0.2053697  1.047293
#   2     extratrees  5.167655  0.1540213  1.103048
#   5     variance    5.187039  0.1710128  1.034382
#   5     extratrees  5.169570  0.1557868  1.122845
#   8     variance    5.633427  0.1247189  1.035171
#   8     extratrees  5.182860  0.1563555  1.137332
# 
# Tuning parameter 'min.node.size' was held constant at a value of 5
# RMSE was used to select the optimal model using the smallest value.
# The final values used for the model were mtry = 2, splitrule =
#  variance and min.node.size = 5.

varimp_mars <- varImp(model.all) 
varimp_mars$importance <- varimp_mars$importance %>% 
  rownames_to_column() %>% 
  mutate(rowname = ifelse(rowname == "dist", "distance to the nearest city",
                          ifelse(rowname == "dist.retail", "distance to the nearest retail centre", 
                                 ifelse(rowname == "London", "distance to London",
                                        ifelse(rowname == "n.nearest.retail.lag", "Nearest retail centre's web density, t-1",
                                               ifelse(rowname == "n.London.lag", "London's web density, t-1",
                                                      ifelse(rowname == "n.l.slag", "spatial and temporal lag of web density",
                                                             ifelse(rowname == "n.nearest.city", "Nearest city's web density, t-1", rowname)))))))) %>% 
  column_to_rownames()

@@

  set.seed(71)
  indices <- CreateSpacetimeFolds(df %>% filter(RGN11CD!=i), 
                                spacevar = "oa11cd",
                                timevar = "year", 
                                k = k)

@@

  #file <- paste0(path, "/outputs/rf/models/oa_rf/regions/region_", as.character(i), ".RData")
  file <- paste0("/hdd/tmp/regions/region_", as.character(i), ".RData")  
  save(model.all, file = file)  # file name = region not in the training data
  rm(model.all)
